Promotors' reports on the first draft of the thesis document.

### 2020-09-09 ChapterConclusion/ChapterConclusion.tex
One before the very last

>>>>

\color{blue}
\subsection*{Continuous Newton method and the error of the Taylor series expansion}

The investigation of the error bounds for the second order Taylor expansion can be conducted using the continuous Newton method. 
This is proposed as future research and the motivation is as follows.

Neuberger's paper \cite{Neuberger07} introduces five theorems that are a consequence of the continuous Newton method.
The Newton method aims to find a zero of the function $F: R^n \rightarrow R^n$ by iterating
\begin{equation} z_{k+1} = z_k - \left(F' (z_k) \right)^{-1} F(z_k) \end{equation}
for $k=0,1,2,\ldots$, and for a given initial $z_0$, assuming that $\left(F' (y) \right)^{-1}$ exists for some $y \in R^n$.
A domain of attraction is the set of all initial values $z_0$ that lead to a root of $F$ after the convergence $z_0, z_1, z_2,\ldots$.
The damped Newton method  
\begin{equation} z_{k+1} = z_k - \delta_k \left(F' (z_k) \right)^{-1} F(z_k) \end{equation}
prevents convergence issues, such as chaotic domains of attraction, with $\delta_1, \delta_2, \ldots \in (0, 1)$. 
The continuous Newton method is a sequence of damped Newton methods. 
To implement the continuous Newton method, select $0<T<1$ and perfom $m$ runs of the damped Newton method with $\delta_k = T/m$, for $k=1,2,\ldots, m$.
The result $z_{k+1}$ is the zero estimation $x_m$, and if the sequence of estimates $x_1, x_2, \ldots$ converges, then it converges to the result of the continuous Newton method.

The objective of the continuous Newton method is to find a function $z: [0, \infty) \rightarrow R^n$, so that
\begin{equation} z(0) = x \in R^n, \quad z' (t) = - \left(F' (z(t)) \right)^{-1} F(z(t)), \quad t \geq 0, \label{eqn:contNewton}\end{equation}
subject to the existence of the limit $u=\mathrm{lim}_{t \rightarrow \infty}{z(t)}$, that satisfies $F(u)=0$.
This implies 
\begin{equation}  F' (z(t)) z' (t) = - F(z(t)) , \end{equation}
that is equivalent to
\begin{equation}  \left(F \circ z\right)' (t) = - F(z(t)) , \end{equation}
from where we get
\begin{equation}  F(z(t)) = e^{-t} F(z(0)) . \end{equation}
Thus, the residual $F(z(t))$ only changes in magnitude and not in direction.
This property is essential for the results of the theorems in \cite{Neuberger07}. 
In particular, Theorem 5 gives the conditions for the continuous Newton method to ensure finding a root of the function.

\begin{customthm}{5} \label{thm:five}
Suppose that each of $H$, $J$, and $K$ is a Banach space, with $H$ compactly embedded in $J$, that $r>0$, and that $G:B_r(0) \rightarrow K$ is continuous as a function on $J$. Suppose also that $g$ belongs to $K$ and that for each $y$ in $b_r(0)$ there is an $h$ in $B_r(0)$, where $b_r(0)$ and $B_r(0)$ are open and closed balls in $H$ of radius $r$ centered at 0, such that
\begin{equation} \lim\limits_{t \to 0+} -{\dfrac{1}{t} \left(G(y+th) - G(y)\right) = g}. \end{equation}
Then, there exists $u$ in $B_r(0)$ such that $G(u)=g$.
\end{customthm}
%$F:B_r(0) \rightarrow K$

The theorem implies that $u$ is a zero of the function  defined as $F(y) = G(y) - g$.  
Since $G$ is a continuous function and if $\left(G' (y) \right)^{-1}$ exists for each $y$ in $B_r(0)$, a metric for the validity of the Theorem (\ref{thm:five}) hypothesis is the inequality $\left\Vert  \left( G' (y) \right)^{-1} g \right\Vert \leq r$.

To link the data-driven step-input estimation problem with this theorem, we can take the system of linear equations $\widetilde{\mathbf{y}} = \widetilde{\mathbf{K}} \bm{\theta}$, then define $G(y) = \widetilde{\mathbf{K}} \bm{\theta}$, and $g = \widetilde{\mathbf{y}}$, so that we have $F(\bm{\theta}) = G(\bm{\theta}) - g$. The derivative of the function $G(\bm{\theta})$ is given by $G'(\bm{\theta}) = \widetilde{\mathbf{K}} \mathbf{I}$. Using the pseudo-inverse of $\widetilde{\mathbf{K}}$, the metric $\left\Vert  \widetilde{\mathbf{K}}^{\dagger} \widetilde{\mathbf{y}} \right\Vert \leq r$ indicates that the estimate $\hat{\bm{\theta}}$ is found inside a ball of radius $r$.

The questions that remain open are how large can $r$ be?, and how is the radius $r$ related to the perturbation noise variance?

\subsection*{The data-driven input estimation methods with missing data}

The implementation of the data-driven step input estimation method can be compromised when one or more step response samples are not properly acquired.
The missing data in the construction of the Hankel matrix may be due to sample transference errors, or due to sensor saturation problems.
The saturation values in the sensor response are not valid for the computation of the step input level because they violate the linearity assumption of the sensor.
The solution to the most powerful unfalsified model identification problem in presence of missing data was provided in \citet{Markovsky16}. 
The methods presented in this work are formulated as a matrix completion with the constraint that the to-be-completed matrix must be a rank deficient Hankel matrix.
The low-rank approximation of Hankel structured matrices with missing data has been studied in \citet{Markovsky13}, where the proposed solution is based on a variable-projection method that requires optimization.
The software for this method, including a link to obtain it, is described in \citet{Markovsky14}. 
This method is applicable in matrix completion and system identification problems and has inspired the development of the signal recovery method in for nuclear magnetic resonance spectroscopy described in \citet{Ying18}, and data compression and feature recognition in \citet{Hou17} based on sparse low-rank matrix approximation.

The solutions of the low-rank approximation with missing data problems are found by iterative optimization and, therefore, require large amount of computational power.
In order to implement data-driven input estimation methods with missing data, there is a need for efficient methods of small complexity $O(n)$ that solve online the optimization problem.
This is an open problem for future research.

	\subsection*{Compressed sensing methods for recovering missing data}

	Compressed sensing is a methodology that is being developed quite fast and that can offer an alternative to perform measurements with missing data. 
	The aim of compressed sensing is the recovery of the missing samples of a signal by computing a sparse coefficient vector, under the assumption that the signal belongs to a generating system of larger order. 
	The missing data estimation problem can be expressed as an overdetermined system of equations with a sparse regression matrix. 
	The review presented in \cite{Eldar2012compressed} provides an introduction of the mathematical modeling of compressed sensing, and gives the conditions for sparse recovery. 
	The recovery algorithms include convex optimization, greedy and combinatorial algorithms, to name a few.

    According to \cite{Rani18}, compressed sensing is applied when we cannot afford buying enough sensors like in non-visible wavelengths applications, when the measurements themselves are expensive as in neutron-scattering imaging, when measurements take a lot of time as in medical imaging, and when a constrain of the measurement is the power consumption.
    A particular field of application for compressed sensing is the recovery of bioelectrical signals, as it is detailed in \cite{Craven15}.

    Another good review of compressed sensing is found in \cite{Duarte11}, where it is explained that the  research community is splitted into a group that studies the problems in which the compressed sensing matrices exhibit structure due to the measurement system modeling, and another group which considers that the signal structured representations not only depend on sparsity but instead assume that the signals belong to continuous time and have infinite-dimensional representations. The solutions provided exploit these structures in the development of the algorithms. Some examples of these structure matrices are described in \cite{Kezhi15} including Gaussian, Toeplitz, Hadamard, and structurally random matrices.

    It might be interesting to compare the solutions of the low-rank approximation methods with compressed sensing in terms of computational complexity and feasibility for online implementations. 
    Therefore, it is proposed as a future research to develop a compressed sensing algorithm, that can estimate directly the input true value by processing the sensor response to the unknown input, and in presence of missing data.

\color{black}

### 2020-09-09 ChapterExperimentalValidation/ChapterExperimentalValidation.tex
One before the very last

\begin{quote}
\vspace{-0.75cm}
\emph{The results of this chapter were published in Quintana Carapia, G., Markovsky, I. Pintelon, R., Csurcsia, P.Z., and Verbeke, D., "Experimental validation of a data-driven step input estimation method for dynamic measurements", IEEE Transactions on Instrumentation and Measurement Journal, \color{blue} vol. 69, no. 7, pp. 4843-4851, July 2020, \color{black} doi: 10.1109/TIM.2019.2951865. \nocite{QuintanaTIM} }\vfill{}
\end{quote}

%\vfill{}

@@

@@

The mean squared error (MSE) of the step input estimate, defined as
\begin{equation} \mathrm{MSE} = b^2 + v,   \end{equation}
where $b$ and $v$ are the bias and the variance of the step input estimate, can be applied to the obtained empirical and predicted results and can be compared to the Cram\'er-Rao lower bound (CRLB).
Figure \ref{fig:MSE_CRLB} shows that $\mathrm{MSE}_\mathrm{e} = b_\mathrm{e}^2 + v_\mathrm{e}$ and $\mathrm{MSE}_{\widetilde{\mathrm{p}}}  = b_{\widetilde{\mathrm{p}}}^2 + v_{\widetilde{\mathrm{p}}}$ have the same proportionality with respect to the measurement noise variance as the bound for an unbiased estimator $\mathrm{CRLB_{ub}}$.
For SNR above 35 dB, $\mathrm{MSE}_\mathrm{e}$ and $\mathrm{MSE}_{\widetilde{\mathrm{p}}}$ are equivalent, and below 35 dB the difference between them is of less than a factor of 10.

We obtained an approximation of the $\mathrm{CRLB}_{\mathrm{b}}$ for our biased estimator using the partial derivative of the bias in expression (\ref{eqn:biasST}).
Figure \label{fig:MSE_CRLB} shows that the bounds for the unbiased and biased estimators are almost equal because the partial derivatives of the bias are negligible w.r.t. 1 in Equation (\ref{eqn:CRLB_EIV}).
By adding the square of the predicted bias to the biased estimator bound $\mathrm{CRLB_{b}}$ we obtain an approximation of the minimum MSE that the biased estimator can achieve.
This minimum MSE is close to the CRLBs for large SNR but the square of the bias causes an increase of the MSE around 35 dB.
The differences between the CRLBs and $\mathrm{MSE}_\mathrm{e}$ and $\mathrm{MSE}_{\widetilde{\mathrm{p}}}$ are of one order of magnitude for large SNR and become small for SNR lower than 40 dB.
This difference is the cost of solving a structured EIV problem with a simple LS method.
%Therefore, for large measurement noise variance, the step input estimation method produces results that are comparable to maximum-likelihood estimation, since $\mathrm{MSE}_{\widetilde{\mathrm{p}}} \approx \mathrm{CRLB_{b}} + b_{\widetilde{\mathrm{p}}}^2$.

\begin{figure}[!htb]
\centering
\includegraphics[width=0.69\columnwidth]{./ChapterExperimentalValidation/fig/Fig_3.pdf}
\caption{\label{fig:MSE_CRLB} The observation instant is fixed at 500 samples.
The bias and the MSEs decrease for large SNR. 
The empirical $\mathrm{MSE_e}$ and the predicted $\mathrm{MSE}_{\widetilde{\mathrm{p}}}$ of the step input estimation are one order of magnitude larger than the Cram\'er-Rao lower bound.
Adding the CRLB for a biased estimator with the predicted bias squared we have the minimum MSE, that grows in the interval [25 dB, 45 dB].
Below 35 dB the difference between $\mathrm{MSE}_{\widetilde{\mathrm{p}}}$ and the CRLB is less than a factor of 10.}
\end{figure}

@@

>>>>

\begin{quote}
%\vspace{-0.75cm}
\emph{The results of this chapter were published in Quintana Carapia, G., Markovsky, I. Pintelon, R., Csurcsia, P.Z., and Verbeke, D., "Experimental validation of a data-driven step input estimation method for dynamic measurements", IEEE Transactions on Instrumentation and Measurement Journal, \color{blue} vol. 69, no. 7, pp. 4843-4851, July 2020, \color{black} doi: 10.1109/TIM.2019.2951865. \nocite{QuintanaTIM} } %\vfill{}
\end{quote}

\vfill{}

@@

\color{blue} 
The simulations described in Chapter 4 were conducted with an illustrative example that is assumed to be a second order weighing sensor.
A weighing sensor was implemented to relate experimental results with the simulations. 
As it will be detailed later, the order of the practical sensor is at least is fifth order. 
Thus, the simulations presented in this section are useful to get insight on the impact of that increasing the model order has on the accuracy of the bias and variance predictions.    
\color{black}

@@

The mean squared error (MSE) of the step input estimate, defined as
\begin{equation} \mathrm{MSE} = b^2 + v,   \end{equation}
where $b$ and $v$ are the bias and the variance of the step input estimate, can be applied to the obtained empirical and predicted results and can be compared to the Cram\'er-Rao lower bound (CRLB).
Figure \ref{fig:MSE_CRLB_str_exp} shows that $\mathrm{MSE}_\mathrm{e} = b_\mathrm{e}^2 + v_\mathrm{e}$ and $\mathrm{MSE}_{\widetilde{\mathrm{p}}}  = b_{\widetilde{\mathrm{p}}}^2 + v_{\widetilde{\mathrm{p}}}$ have the same proportionality with respect to the measurement noise variance as the bound for an unbiased estimator $\mathrm{CRLB_{ub}}$.
For SNR above 35 dB, $\mathrm{MSE}_\mathrm{e}$ and $\mathrm{MSE}_{\widetilde{\mathrm{p}}}$ are equivalent, and below 35 dB the difference between them is of less than a factor of 10.

We obtained an approximation of the $\mathrm{CRLB}_{\mathrm{b}}$ for our biased estimator using the partial derivative of the bias in expression (\ref{eqn:biasST}).
Figure \ref{fig:MSE_CRLB_str_exp} shows that the bounds for the unbiased and biased estimators are almost equal because the partial derivatives of the bias are negligible w.r.t. 1 in Equation (\ref{eqn:CRLB_EIV}).
By adding the square of the predicted bias to the biased estimator bound $\mathrm{CRLB_{b}}$ we obtain an approximation of the minimum MSE that the biased estimator can achieve.
This minimum MSE is close to the CRLBs for large SNR but the square of the bias causes an increase of the MSE around 35 dB.
The differences between the CRLBs and $\mathrm{MSE}_\mathrm{e}$ and $\mathrm{MSE}_{\widetilde{\mathrm{p}}}$ are of one order of magnitude for large SNR and become small for SNR lower than 40 dB.
This difference is the cost of solving a structured EIV problem with a simple LS method.
%Therefore, for large measurement noise variance, the step input estimation method produces results that are comparable to maximum-likelihood estimation, since $\mathrm{MSE}_{\widetilde{\mathrm{p}}} \approx \mathrm{CRLB_{b}} + b_{\widetilde{\mathrm{p}}}^2$.

\begin{figure}[!htb]
\centering
\includegraphics[width=0.69\columnwidth]{./ChapterExperimentalValidation/fig/Fig_3.pdf}
\caption{\label{fig:MSE_CRLB_str_exp} The observation instant is fixed at 500 samples.
The bias and the MSEs decrease for large SNR. 
The empirical $\mathrm{MSE_e}$ and the predicted $\mathrm{MSE}_{\widetilde{\mathrm{p}}}$ of the step input estimation are one order of magnitude larger than the Cram\'er-Rao lower bound.
Adding the CRLB for a biased estimator with the predicted bias squared we have the minimum MSE, that grows in the interval [25 dB, 45 dB].
Below 35 dB the difference between $\mathrm{MSE}_{\widetilde{\mathrm{p}}}$ and the CRLB is less than a factor of 10.}
\end{figure}

@@

\color{blue}

	There exist methods for estimating the performance of different models that are useful to choose among the models.
	The model selection methods can be analytical, such as the Akaike information criterion (AIC) introduced in the paper of \citet{Akaike74}, the minimum description length (MDL) iniciated by the work of \citet{Rissanen78}, or numerical such as crossvalidation.
	The AIC is defined for a class of models fit by maximum likelihood. 
	For the linear regression problem with Gaussian perturbation (3.20), the AIC is given by
	\begin{equation} \mathrm{AIC} = 2 \left( n+1 \right) - \mathrm{ln} \left( \widehat{L} \right) \label{eqn:AIC} \end{equation}
	where $n+1$ is the number of estimated parameters, and $\widehat{L}$ is the maximum value of the likelihood function of the model.
	The AIC rewards the goodness of fit by considering the maximum likelihood value, and adds a penalty term that increases the cost with respect to the number of parameters to prevent overfitting.
	In the opinion of \citet{McElreath20}, the AIC is reliable when the priors overwhelmed by the likelihood, the posterior distribution is approximately multivariate Gaussian, and the sample size is much larger than the number of parameters.
	Using this criterion for model selection is simple, \citet{Hastie09} suggest just picking the model that provides the smallest AIC value.

	The MDL approach is motivated from an optimal coding viewpoint, aiming to choose the most parsimonious model, that, according to \citet{Hastie09}, is the shortest model that can encode the data for transmision of messages.
	If we have a model with parameters $\theta$, and data $(K, y)$, with the conditional probability of the outputs $\mathrm{P}(y | \theta, K)$, and assuming the receiver knows the inputs and we transmit the outputs, then the message length required to transmit the outputs is
    \begin{equation} \mathrm{length} = - \mathrm{ln} \left( \mathrm{P}( \theta ) \right) - \mathrm{ln} \left( \mathrm{P}(y | \theta, K) \right) . \label{eqn:MDL} \end{equation}
    The length is the sum of the code length required for transmitting the model parameters with the code length required for transmitting the discrepancy between the model and the actual target values. 
    The recommended model is the one that minimizes (\ref{eqn:MDL}).
	\citet{McElreath20} arguments that model fitting is a form of data compression, where parameters explain the relationships among the data. 
	The parameters compress the data into a simpler form, with some loss of information about the sample, and 
	can be used to generate new data by decompression.
	Learning about the data requires using a simpler model that achieves some compression, but not too much.
    \color{black}

### 2020-09-09 ChapterIntroduction/ChapterIntroduction.tex
One before the very last

\begin{figure}[!htbp]
\centering
\includegraphics[width=0.69\columnwidth]{./ChapterIntroduction/fig/Fig_2.pdf} 
\caption{ \label{fig:thermometer} 
\color{blue} This simulation shows an example of the temperature change experimented by a thermometer, with $k_{\mathrm{T}}=0.5 s^{-1}$, caused by a step input of 67 $^{\circ} \mathrm{C}$  when the thermometer was initially at 23 $^{\circ} \mathrm{C}$. In this example, we should wait 10 s to have an estimation of the input with a relative error lower than 0.5 \%. \color{black} }
\end{figure}

\begin{figure}[!htbp]
\centering
\includegraphics[width=0.69\columnwidth]{./ChapterIntroduction/fig/Fig_3.pdf} 
\caption{ \label{fig:scale} 
\color{blue} This simulation illustrates the displacement observed when a constant input $u=1 \mathrm{kg}$ is applied to a weighing sensor in equilibrium. The sensor parameters are $k_{\mathrm{d}} = 5 \ \mathrm{N \ s/m}$, $m = 5 \ \mathrm{g}$, and $k_{\mathrm{s}} = 10250 \ \mathrm{N/m}$. In this conditions, the user has to wait more that 2 s to observe a stabilization on the scale and to estimate the mass from reading the final value of the displacement with a relative error smaller than 0.72 \%. \color{black}  }
\end{figure}

@@

\begin{figure}[!htbp]
\centering
\includegraphics[width=0.69\columnwidth]{./ChapterIntroduction/fig/Fig_4.pdf} 
\caption{ \label{fig:force} 
\color{blue} The simulation of a force sensor based on a double mass-spring system in cascade also shows that the stabilization after a unit step input is not instantaneous. In the plot we infer that the user has to wait 5 s to estimate the force from the sensor response that is 0.25 \% far from its value in equilibrium. \color{black}  }
\end{figure}

The handbook of sensors written by \citet{Fraden16Book} describes a pressure sensor that detects sudden changes in the pressure inside a room.
These changes can be caused by the opening of a door or by the movement of a person.
The principle of operation of detecting minute pressure gradients is the flowmeter that compares the pressure at both outlets of a tube.   
An experimental identification of such a flowmeter is described in \citet{Wiklund02}, where a second order model is proposed to explain the dynamics with the transfer function
\begin{equation} H(s) = \dfrac{ e^{-\tau_{delay}s}}{\left( 1 +  \tau_1 s \right) \left( 1 +  \tau_2 s \right)} 
\end{equation}  
where $\tau_{delay} = 0.07$ s is the dead time, and $\tau_{1} = 0.01$ s and $\tau_{2} = 1.6$ s are two time constants.
The step response simulation of this dynamic model results in the sensor step response shown in Figure \ref{fig:pressure}.  
We have a response to a unit step function that is mostly and increasing exponential that requires 7 s approximately to reach a value that is 98.7 \% ot its maximum value.

@@

\section{Original contributions}

@@

This thesis describes research work initiated after the data-driven step input estimation method proposed by \citet{Markovsky15cep}.
This method aims for metrology applications but it lacked an uncertainty assessment, and, therefore, its appropriateness was questionable.
The first research work conducted was a statistical analysis to elucidate the bias and the covariance of the estimate that the data-driven step input estimation method provides.
After obtaining these statistical moments, the estimation uncertainty was assessed, and thus, the effectiveness of the method was appraised.
The second research work was a series of experiments conducted to validate the data-driven step input estimation method in real-life applications.
The experiments were realizations of the step input excitation using temperature and mass sensors.
The step input estimation method showed robustness when the measurement noise is not Gaussian and white, as it was assumed in the theoretical analyses.
The third research work conducted was the estimation of affine input parameters following the data-driven approach for real-time applications. 
The data-driven method uses exponential weighing in the recursive least-squares solution of a structured errors-in-variables problem, to give preference to recent samples over the older samples.
The methodology and performance of the data-driven affine input estimation method were compared to those of a maximum-likelihood method.

@@

@@

Instead of using a total least-squares (TLS) solution of the structured EIV problem, the step input estimation method uses the recursive least-squares solution (RLS). 
In the review conducted by \citet{Markovsky07overview} it is shown that the TLS solution of unstructured EIV problems is consistent when the perturbations have zero mean with a given positive definite covariance, and
the TLS solution is equivalent to the maximum likelihood (ML) solution when the disturbances of the EIV problems are i.i.d. normally distributed . 
However, the studies of structured EIV problems presented in \citet{VanHuffel07TLSeditorial} demonstrate that the TLS estimator does not give general results since each structure requires a specific treatment. 
Moreover, \citet{Rhode14recursive} found that the ML estimator leads to non-convex optimization when applied to structured EIV problems, and the global optimum is not guaranteed.
Also, the computational complexity of TLS and ML inhibits their real-time implementation.
Thus, the least-squares (LS) estimator is a suboptimal but simple solution to structured EIV problems that admits a recursive form (RLS) suitable for real-time implementation.

@@

@@

@@

>>>>

\begin{figure}[!htbp]
\centering
\includegraphics[width=0.69\columnwidth]{./ChapterIntroduction/fig/Fig_2.pdf} 
\caption{ \label{fig:thermometer} 
\color{blue} This simulation shows an example of the temperature change experimented by a thermometer, with $k_{\mathrm{T}}=0.5 s^{-1}$, caused by a step input of 67 $^{\circ} \mathrm{C}$  when the thermometer was initially at 23 $^{\circ} \mathrm{C}$. In this example, we should wait 10 s to have an estimation of the input with a relative error lower than 0.5\%. \color{black} }
\end{figure}

\begin{figure}[!htbp]
\centering
\includegraphics[width=0.69\columnwidth]{./ChapterIntroduction/fig/Fig_3.pdf} 
\caption{ \label{fig:scale} 
\color{blue} This simulation illustrates the displacement observed when a constant input $u=1 \mathrm{kg}$ is applied to a weighing sensor in equilibrium. The sensor parameters are $k_{\mathrm{d}} = 5 \ \mathrm{N \ s/m}$, $m = 5 \ \mathrm{g}$, and $k_{\mathrm{s}} = 10250 \ \mathrm{N/m}$. In this conditions, the user has to wait more that 2 s to observe a stabilization on the scale and to estimate the mass from reading the final value of the displacement with a relative error smaller than 0.72\%. \color{black}  }
\end{figure}

@@

\begin{figure}[!htbp]
\centering
\includegraphics[width=0.69\columnwidth]{./ChapterIntroduction/fig/Fig_4.pdf} 
\caption{ \label{fig:force} 
\color{blue} The simulation of a force sensor based on a double mass-spring system in cascade also shows that the stabilization after a unit step input is not instantaneous. In the plot we infer that the user has to wait 5 s to estimate the force from the sensor response that is 0.25\% far from its value in equilibrium. \color{black}  }
\end{figure}

The handbook of sensors written by \citet{Fraden16Book} describes a pressure sensor that detects sudden changes in the pressure inside a room.
These changes can be caused by the opening of a door or by the movement of a person.
The principle of operation of detecting minute pressure gradients is the flowmeter that compares the pressure at both outlets of a tube.   
An experimental identification of such a flowmeter is described in \citet{Wiklund02}, where a second order model is proposed to explain the dynamics with the transfer function
\begin{equation} H(s) = \dfrac{ e^{-\tau_{delay}s}}{\left( 1 +  \tau_1 s \right) \left( 1 +  \tau_2 s \right)} 
\end{equation}  
where $\tau_{delay} = 0.07$ s is the dead time, and $\tau_{1} = 0.01$ s and $\tau_{2} = 1.6$ s are two time constants.
The step response simulation of this dynamic model results in the sensor step response shown in Figure \ref{fig:pressure}.  
We have a response to a unit step function that is mostly and increasing exponential that requires 7 s approximately to reach a value that is 98.7\% of its maximum value.

@@

\section{Original contributions \color{blue}and organization of the thesis\color{black}}

@@

This thesis describes research work initiated after the data-driven step input estimation method proposed by \citet{Markovsky15cep}.
This method aims for metrology applications but it lacked an uncertainty assessment, and, therefore, its appropriateness was questionable.
\color{blue}A description of the data-driven step input estimation method is presented in Chapter 3\color{black}.

@@

\color{blue}Chapter 4 describes \color{black} the first research work conducted that is a statistical analysis of the data-driven step input estimation method, aiming to elucidate the bias and the covariance of the input estimation.
After obtaining these statistical moments, the estimation uncertainty was assessed, and thus, the effectiveness of the method was appraised.

@@

Instead of using a total least-squares (TLS), \color{blue} or an intrumental variables \color{black}, solution of the structured EIV problem, the step input estimation method uses the recursive least-squares solution (RLS). 
In the review conducted by \citet{Markovsky07overview} it is shown that the TLS solution of unstructured EIV problems is consistent when the perturbations have zero mean with a given positive definite covariance, and
the TLS solution is equivalent to the maximum likelihood (ML) solution when the disturbances of the EIV problems are i.i.d. normally distributed. 
\color{blue} The classical TLS solution is obtained using singular value decomposition of the augmented matrix built from the regression matrix and the regressor as additional columns\color{black}.
However, the studies presented in \citet{VanHuffel07TLSeditorial} demonstrate that the structured TLS solutions cannot be generalized since each structure in an EIV problem requires a specific treatment. 
Moreover, the ML estimator of structured EIV problems leads to non-convex optimization, and the global optimum is not guaranteed \cite{Beck10}.
\color{blue} On the other hand, the instrumental variable (IV) solution of EIV problems use an additional variable.
The additional variable is an instrument with the purpose of removing the bias of the estimate.
To do this, the construction of the instrumental variable has to be done in such a way that it is uncorrelated with the perturbations and correlated with the regressor, as it is described in \citet{Soderstrom18}. 
\color{black} 
Unfortunately, the computational complexity of the TLS, ML, and IV solutions inhibits their real-time implementation.

\color{blue} 
The solution of the step input estimation problem is preferred to be online, since its applications are in the field of metrology.
In the literature there are references for online implementations of TLS, ML or IV methods, which are are mainly recursive versions of the solution methods.
These recursive versions are developed to solve EIV problems for system identification and fault detection.
We have, for example, the generalized recursive TLS solutions introduced in \citet{Rhode14recursive} and \citet{Rhode14recursivecov}, and the recursive IV solutions developed in \citet{Djouambi12}, \citet{Shang16} and \citet{Gil15}.
These methods cannot be used as off-the-shelf solutions to the step input estimation EIV problem because they need to be adapted to the specific requirements of the problem.
One is that the sensor output is the only measured signal, since the input is to-be estimated, and the solution should be simpler, preferable with a linear complexity $O(n)$, instead of the cubic $O(n^3)$ or cuadratic $O(n^2)$ complexities of the mentioned recursive methods, where $n$ is the sensor order.
The proposal of the data-driven step input estimation method includes the recursive least-squares (RLS) solution.
Compared with recursive TLS or recursive IV solutions, RLS is a suboptimal but more simple solution to the structured EIV problem with lower computational complexity,  suitable for real-time implementation.
\color{black}

@@

\color{blue}Chapter 5 describes \color{black} the second research work that consisted in a series of experiments conducted to validate the data-driven step input estimation method in real-life applications.
The experiments were realizations of the step input excitation using a mass sensor.
The step input estimation method showed robustness when the measurement noise is not Gaussian and white, as it was assumed in the theoretical analyses.

@@

\color{blue}Chapter 6 describes \color{black} the third research work conducted to estimate the parameters of an affine input, following the data-driven approach for real-time applications. 
The data-driven method uses exponential weighing in the recursive least-squares solution of a structured errors-in-variables problem, to give preference to recent samples over the older samples.
The methodology and performance of the data-driven affine input estimation method were compared to those of a maximum-likelihood method.

@@

\subsection{\color{blue}Main findings\color{black}}
\color{blue}Chapter 7 concludes this thesis work and summarizes the main results obtained in the conducted research.

The data-driven step input estimation method reduces the input estimation time compared with formulations based on models of the sensor.
This is because the estimation method does not require to invert the sensor dynamics.
Instead, an algorithm in a digital signal processor finds directly the step input value from the sensor response.  
Since the method is not rescricted to a particular sensor model, the method can process the step response of any sensor, and provides the input estimation in different applications. 
Moreover, the method requires relatively low computational power to process the signal of the sensor.
For these reasons, the data-driven method is suitable for real-time measurements.

After studying the statistical properties of the step input estimation, we know the accuracy (bias) and the precision (variance) of the results.
This knowlewdge allows one to understand what is the uncertainty on the step input value that the method estimates.
We have derived formulas to quantify the bias and the variance of the step input estimation.
With these formulas we can foresee the bias and standard deviation we will obtain with the method given the number of samples processed and the level of the measurement noise.

The implementation of the bias and variance calculation is illustrated with a real life sensor.
A weighing system, that uses a load cell sensor, is used to generate real life sensor step responses. 
The measurement noise observed from this weighing sensor is not Gaussian white noise.
The study of the estimation method performance with real data suggests what is the method order required to set a proper implementation.

It was explored the applicability of data-driven methods to estimate an affine input from the sensor response.
This input model is observed in measurements where the input changes with a constant rate. 
The signal processing algorithm, studied in the previous stages of the research, was extended to solve the affine input estimation problem.   
The performance results of the extended method is compared with a maximum-likelihood method that solves the affine input estimation problem using optimization.

We learnt that the data-driven input estimation methods are simple-to-use methods that run on digital signal processors with few computational resources and estimate the unknown inputs in real time, where the estimation uncertainty can be deduced from the amount of data processed and measurement noise variance.

### 2020-09-09 ChapterPreliminaries/ChapterPreliminaries.tex
One before the very last

%\vfill{}

@@

@@

>>>>

%\vfill{}
\color{blue}
The dynamic systems in the real world are non linear. 
Under certain intervals of operation, the linear approximations can explain the dynamics of the systems. 
The linear and time invariant systems have been studied in depth and the linear theory is a solid collection of tools to analyze and explain the behaviour of dynamic systems in a wide range of applications. 
On the other hand, the study of non-linear systems is more complex and strongly depends on the type of non-linearity that each particular systems exhibit. 
When the non-linear modeling does not apport a big amount of benefits into the application study, such as significative accuracy improvement or supporting wide ranges of operation, then is better to resort to the utilization of LTI representations. 
A sensor is a dynamic system whose operation range is preferred to be linear, to facilitate the repeatability and reproducibility of the measurements.
This means that every time the sensor must deliver identical responses for the same input excitations.   
\color{black}

@@

\color{blue} 
The process noise is the representation of the uncertainty in the system state due to unmodeled perturbations that affect the state evolution. 
The process noise is modeled as an additional input to the system excitation. 
For a linear time invariant (LTI) system, the response to a sum of two input excitations is the sum of the two responses that correspond to the individual inputs.
It is demonstrated in \citet{Gubner06} and \citet{Smith11} that the response of an LTI system to a white noise input is colored noise, since the power spectral density of the response is not constant, but instead proportional to the squared magnitude of the system impulse response.
Nevertheless, in the bandwidth operation of a properly designed measurement system, processing white noise results in approximately white noise output. 
Moreover, in the present work the measurement noise exists only in the system response signal, because the input is not direcly observed, and the system is not subject to external perturbation. 
Under these conditions, the mean and variance of the measurement noise $\epsilon$ describe implicitly the process noise entering the system. 
\color{black}

@@

\color{blue}
{\bfseries Note.} The formulation of the data-driven step input estimation method is based on the behavioral theory introduced in \citet{Willems86I}.
A behaviour of a dynamical system is any set of input-output trajectories that belongs to the system. 
It is demonstrated in \citet{Willems05} that when an input-output trajectory is persistently exciting enough, a Hankel matrix constructed from this trajectory represents the most powerful unfalsified model.
This model is unfalsified because it describes exactly all possible trayectories of the system, and is the most powerful because it is the system representation with the minimal complexity.     
The rank of the Hankel matrix constructed from a trajectory is a metric used to find the order of persistency of the trajectory.
The rank-deficient Hankel matrix with the smaller number of block rows indicates the trajectory order of persistency. %
When the system is excited by a step input, the system is equivalent to an autonomous augmented system, and a model for it can be obtained from the Hankel matrix constructed only from the step response.

\color{black}
